\chapter{Reference: Implementation, Data Tables, Register Status}
\label{ch:anhang-referenz}

\section*{The reference implementation}

All enumerations and verifications in this book rest on a small,
deliberately unoptimized Python implementation
(\texttt{horimetrik\_reference.py}) with self-tests: encoding and
decoding, structure polynomials, both extensions, both
compactifications, structure arithmetic. It is complemented by
dated experiment scripts whose results are kept in the working
register with a status (proved / conjectured / refuted). The
project's guiding principle: enumeration produces conjectures and
counterexamples, never theorems -- and every claimed bound must
survive constructed extreme cases, not merely random samples.

\section*{The nine counting sequences}

\begin{center}
\footnotesize
\begin{tabular}{lll}
\toprule
Sequence & Initial values & Status\\
\midrule
Jump positions $m_g$ ($g{\ge}1$) & $3,15,84,533,3883,31998,294875,\ldots$ &
Depth law, asymptotics proved\\
$\beta(r,r)$ ($r{\ge}1$) & $1,6,22,87,407,2473,19527,\ldots$ & Asymptotics proved\\
Jumpers $D(n)$ ($n{\ge}3$) & $1,8,60,360,2520,21168,211680,2268000$ &
Digit formula proved\\
Non-jumpers $N(n)$ ($n{\ge}3$) & $5,16,60,360,2520,19152,151200,\ldots$ &
Digit formula proved\\
Fixed points $F(n)$ ($n{\ge}2$) & $4,17,88,540,3960,32760$ & $=n\cdot n!-D(n)$ (proved)\\
Single-digit factorials & $1,2,5,7,11,23,29,39,59,119,\ldots$ &
fully characterized\\
$\max\Omega$ ($n{\ge}1$) & $0,1,6,43,351,2873,26443,270291$ & $\Omega\ge0$
proved\\
$\#\{\Omega=0\}$ ($n{\ge}1$) & $2,5,13,23,61,111,274,564$ &
Equality criterion proved\\
$\Gamma_\circ(r,r)$ ($r{\ge}1$) & $1,4,23,6541,477149751,\ldots$ &
Extremal principle proved\\
\bottomrule
\end{tabular}
\end{center}

None of these sequences was in the OEIS at the time of writing;
the submission is prepared (A-numbers will be added here).

\section*{Confirmed predictions}

\begin{enumerate}
\item Depth $-3$ of the commutator first near shell $100$
      (exactly: 84 as a jump position, confirmed in the range up
      to 200).
\item $m_6=31998$ against a predicted $\approx31930$.
\item $m_7=294875$ -- exactly as predicted.
\item Single-digit factorials at $n=29,39,59,119$ and
      $143,179,239,359,719$ from the (by now complete)
      characterization $n=(i{+}1)!/d-1$ -- all nine confirmed.
\end{enumerate}

\section*{Register status at this version's editorial deadline}

44 proved theorems -- among them the limit-space theorems of the
three staircases, the two dual semirings, the duality theorem and
the side-fidelity theorem, the compactification monoid, the depth
law, the digit formula, and the interchange positivity --, plus
9 documented dead ends, 3 refuted conjectures of our own,
9 previously unknown counting sequences, and 4 confirmed groups of
predictions (12 individual hits, listed above). The current
version of the register, conjectures, counterexamples, and
changelog accompanies the project's working files.
